---
title: "KDE Stuff"
format: pdf
---

## Average of KDEs

Is the average of several KDEs the same as the KDE of the whole data set? Suppose we have $\mathcal{X}_1 = X_{11}, \dots, X_{1N}$ and $\mathcal{X}_2 =X_{21}, \dots, X_{2N}$. The individual estimates of the densities for $\mathcal{X}_1$ and $\mathcal{X}_2$ are

$$
\hat{f}^{(1)}(x) = \frac1{Nh}\sum_{i=1}^NK(\frac{x - X_{1j}}{h})
$$
and

$$
\hat{f}^{(2)}(x) = \frac1{Nh}\sum_{j=1}^NK(\frac{x - X_{2j}}{h}),
$$
respectively. The KDE of $\mathcal{X}_1$ and $\mathcal{X}_2$ together would be

$$
\hat{f}(x) = \frac1{2Nh}\sum_{i=1}^2\sum_{j=1}^NK(\frac{x - X_{ij}}{h}),
$$

If we take the average of the two estimators, then we have  

\begin{equation}
\begin{split}
\hat{f}^{(1, 2)}(x)  & = \frac12 (\hat{f}^{(1)}(x) + \hat{f}^{(2)}(x))  \\
 & = \frac12 \left[\frac1{Nh}\left(\sum_{j=1}^NK(\frac{x - X_{1j}}{h}) + \sum_{j=1}^NK(\frac{x - X_{1j}}{h})\right)\right] \\
 & = \hat{f}(x).
\end{split}
\end{equation}

Am I missing something? Let's check an example. 


```{r}
#| include: true
set.seed(9283)
df <- dplyr::tibble(u = runif(100),
                    xc1 = rnorm(100, m = 0, s = 1),
                    xc2 = rnorm(100, m = 3, s = 1),
                    x1 = ifelse(u <= .3, xc1, xc2),
                    x2 = rnorm(100, m = 0, s = 1))
d1 <- density(df$x1, bw = .5, from = -3.5, to = 6.5)
d2 <- density(df$x2, bw = .5, from = -3.5, to = 6.5)
d3 <- density(c(df$x1, df$x2), bw = .5, from = -3.5, to = 6.5)
df_dens <- dplyr::tibble(x = c(d1$x, d2$x, d3$x, d1$x),
                         y = c(d1$y, d2$y, d3$y, (d2$y + d1$y)/2),
                         density = rep(c("Mixture", "N(3, 1)", "Combined", "Average"), 
                                       each = 512))
```
```{r}
#| echo: false
#| fig-cap: "Densities of individual points."

p <- ggplot2::ggplot(data = df_dens, 
                     ggplot2::aes(x = x, y = y, col = density))
p + ggplot2::geom_line() +
  ggplot2::scale_color_brewer(palette = "Set1") +
  ggplot2::theme_bw()
```

```{r}
#| echo: false
#| fig-cap: "Densities of individual points (facetted)."

p <- ggplot2::ggplot(data = df_dens, 
                     ggplot2::aes(x = x, y = y, col = density))
p + ggplot2::geom_line() +
  ggplot2::facet_wrap(~density, nrow = 2) +
  ggplot2::scale_color_brewer(palette = "Set1") +
  ggplot2::theme_bw()
```
